---
title: "cell Atlas or Human Cell Landscape Assignments"
author: "Anthony Hung"
date: "2020-01-24"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Introduction

# Cell atlas from Mengie Chen's lab (UChicago)
Compute correlation between reference set of samples and individual cells (input is seurat object.)
Load correlation functions

```{r load functions}
library(Seurat)
library(SingleCellExperiment)
library(Matrix.utils)

#load correlation functions
source("code/cellAtlas_assignment.R")

#load single-cell data from pilot
#ANT1.2 <- readRDS("data/ANT1_2.rds")
SCT_integrated <- readRDS("data/SCT_ANT12_integrated.rds")
```

## Apply functions to single cell data to label them with assigned cell type

```{r assign metadata}
#Compute the correlation matrix use a seurat object (COUNTS) and cell line matrix as input
DefaultAssay(SCT_integrated) <- "RNA"
cor.mat <- ComputeCorMat(SCT_integrated,cell.line)
head(cor.mat)

#Add a metadata column "clusteredCellAtlas" to the object
SCT_integrated <- AssignClusteredLabel(SCT_integrated,cor.mat, K=20, topCorrelation=5)

#Table of assigned cell labels
table(SCT_integrated$clusteredCellAtlas)
```

## Exploring different values of K (clusters of cellatlas populations)

Out of interest, I look at how trying different resolutions of the cellatlas data affect assignments of cell types

```{r try K}
for(kval in 17:57){
     print(kval)
     print(table(AssignClusteredLabel(SCT_integrated,cor.mat, K=kval, topCorrelation=5)$clusteredCellAtlas))
}

```


## Crude analyses of correlation matrix

```{r crude}
# see which cell lines are highest correlated (sum over all cells) in the data with the summed values
colSums(cor.mat)[colSums(cor.mat) > 800]

```

# Human cell Landscape (http://bis.zju.edu.cn/HCL/blast.html)

This is a similar method, but uses DGE objects and a different reference set (the reference set comes from single cell rather than bulk data). Computes correlations between cells in a test sample and the reference set samples and assigns best correlation identity.

```{r HCL prep}
#This require devtools  
#install.packages('devtools')
#library(devtools)
# scHCL requires ggplot2/reshape2/plotly/shiny/shinythemes/shiny
#install_github("ggjlab/scHCL")
library(scHCL)
library(Seurat)
library(scran)
```

Convert to DGE (Keep all genes and use raw unnormalized counts).

```{r DGE}
SCT_integrated <- readRDS("data/SCT_ANT12_integrated.rds")
metadata <- SCT_integrated@meta.data # keep metadata handy
sce <- as.SingleCellExperiment(SCT_integrated, assay="RNA")
dge<- convertTo(sce, type= "edgeR")
remove(sce)
dge$samples<- cbind(dge$samples, metadata)
```

Run scHCL and visualize
No chondrocytes in the data, so maybe it could be added?

```{r hcl_result}
# hcl_result <- scHCL(scdata = dge$counts, numbers_plot = 3)
# scHCL_vis(hcl_result)
```






# Run HCA correlation analysis on the data integrated across individual
Goal is to see whether the cluster 2 coincides with the cells that are more highly correlated with chondrocytes/osteoblasts

```{r individual}
SCT.integrated_individual <- readRDS("data/SCT_integrated_by_individual.rds")

#psuedobulk
data <- SCT.integrated_individual@assays$SCT@counts
metadata <- SCT.integrated_individual@meta.data

# Set up metadata as desired for aggregation and DE analysis
metadata$seurat_clusters <- factor(metadata$seurat_clusters)
metadata$labels <- factor(metadata$labels)

# Create single cell experiment object
sce <- SingleCellExperiment(assays = list(counts = data), 
                           colData = metadata)

# Identify groups for aggregation of counts
groups <- colData(sce)[, c("labels", "seurat_clusters")]

#pseudobulk
pb <- aggregate.Matrix(t(counts(sce)), 
                       groupings = groups, fun = "sum") 
pb_matrix <- as.matrix(pb)
SCT.integrated.pseudobulk <- CreateSeuratObject(t(pb_matrix))
SCT.integrated.pseudobulk$cluster_ind <- colnames(SCT.integrated.pseudobulk@assays$RNA@counts)
```

## Apply functions to single cell data to label them with assigned cell type

```{r assign metadata individual}
#Compute the correlation matrix use a seurat object (COUNTS) and cell line matrix as input
DefaultAssay(SCT.integrated_individual) <- "SCT"
cor.mat <- ComputeCorMat(SCT.integrated_individual,cell.line)
head(cor.mat)

#Add a metadata column "clusteredCellAtlas" to the object
SCT.integrated_individual <- AssignClusteredLabel(SCT.integrated_individual,cor.mat, K=20, topCorrelation=5)

#Table of assigned cell labels (and by cluster assignment)
table(SCT.integrated_individual$clusteredCellAtlas)
table(SCT.integrated_individual$clusteredCellAtlas, SCT.integrated_individual$seurat_clusters)
```

## Exploring different values of K (clusters of cellatlas populations)

Out of interest, I look at how trying different resolutions of the cellatlas data affect assignments of cell types

```{r try K individual}
for(kval in 17:57){
     print(kval)
     print(table(AssignClusteredLabel(SCT.integrated_individual,cor.mat, K=kval, topCorrelation=5)$clusteredCellAtlas, SCT.integrated_individual$seurat_clusters))
}
```



## Apply functions to single cell data to label them with assigned cell type (pseudobulked)

```{r assign metadata individual pb}
#Compute the correlation matrix use a seurat object (COUNTS) and cell line matrix as input
DefaultAssay(SCT.integrated.pseudobulk) <- "RNA"
cor.mat <- ComputeCorMat(SCT.integrated.pseudobulk,cell.line)
head(cor.mat)

#Add a metadata column "clusteredCellAtlas" to the object
SCT.integrated.pseudobulk <- AssignClusteredLabel(SCT.integrated.pseudobulk,cor.mat, K=20, topCorrelation=5)

#Table of assigned cell labels (and by cluster assignment)
table(SCT.integrated.pseudobulk$clusteredCellAtlas)
table(SCT.integrated.pseudobulk$clusteredCellAtlas, SCT.integrated.pseudobulk$cluster_ind)
```

## Exploring different values of K (clusters of cellatlas populations)

Out of interest, I look at how trying different resolutions of the cellatlas data affect assignments of cell types

```{r try K individual pb}
for(kval in 17:57){
     print(kval)
     print(table(AssignClusteredLabel(SCT.integrated.pseudobulk,cor.mat, K=kval, topCorrelation=5)$clusteredCellAtlas, SCT.integrated.pseudobulk$cluster_ind))
}
```
