---
title: "DE_analysis"
author: "Anthony Hung"
date: "2019-12-16"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

```{r}

```








From jdblischak's code:


```{r chunk-options}
library("knitr")
opts_chunk$set(cache = FALSE)
```

Initial limma analysis.

## Setup

```{r packages, message=FALSE}
library("limma")
library("edgeR")
library("dplyr")
library("tidyr")
library("ashr")
library("ggplot2")
library("cowplot")
theme_set(theme_bw(base_size = 14))
theme_update(panel.grid.minor.x = element_blank(),
             panel.grid.minor.y = element_blank(),
             panel.grid.major.x = element_blank(),
             panel.grid.major.y = element_blank())
library("VennDiagram")
```

Input data.

```{r input-counts}
counts <- read.delim("../data/counts.txt", check.names = FALSE, row.names = 1)
anno <- read.delim("../data/experiment-info.txt", stringsAsFactors = FALSE)
rownames(anno) <- anno$id
anno <- anno %>% select(-id, -tube_ind)
stopifnot(colnames(counts) == rownames(anno))
```

## Model with limma

```{r design-matrix}
anno$status <- factor(anno$status, levels = c("contact", "tb"))
anno$treatment <- factor(anno$treatment, levels = c("none", "infected"))
design <- model.matrix(~treatment + status + status:treatment,
                       data = anno)
colnames(design)[1] <- "Intercept"
colnames(design) <- gsub("treatment", "", colnames(design))
colnames(design) <- gsub("status", "", colnames(design))
colnames(design) <- gsub(":", ".", colnames(design))
colnames(design)
```

```{r fit-model, fig.width=8}
y <- DGEList(counts)
y <- calcNormFactors(y)
v <- voom(y, design)
# Model individual as a random effect
corfit <- duplicateCorrelation(v, design, block = anno$individual)
corfit$consensus
v <- voom(y, design, block = anno$individual, correlation = corfit$consensus)
fit <- lmFit(v, design, block = anno$individual, correlation = corfit$consensus)
fit <- eBayes(fit)
```

## Assess model results

```{r model-results}
top_treatment <- topTable(fit, coef = "infected")
top_treatment
results_treatment <- topTable(fit, coef = "infected",
                              number = nrow(counts), sort = "none")
ma_treatment <- ggplot(data.frame(Amean = fit$Amean, logFC = fit$coef[, "infected"]),
                       aes(x = Amean, y = logFC)) +
  geom_point() +
  labs(x = "Average expression level", y = "Log fold change",
       title = "Treatment effect")
ma_treatment
hist_treatment <- ggplot(results_treatment, aes(x = P.Value)) +
  geom_histogram(binwidth = 0.01) +
  labs(x = "p-value", y = "Number of genes", title = "Treatment effect")
hist_treatment
top_status <- topTable(fit, coef = "tb")
top_status
results_status <- topTable(fit, coef = "tb",
                           number = nrow(counts), sort = "none")
ma_status <- ggplot(data.frame(Amean = fit$Amean, logFC = fit$coef[, "tb"]),
                       aes(x = Amean, y = logFC)) +
  geom_point() +
  labs(x = "Average expression level", y = "Log fold change",
       title = "Status effect")
ma_status
hist_status <- ggplot(results_status, aes(x = P.Value)) +
  geom_histogram(binwidth = 0.01) +
  labs(x = "p-value", y = "Number of genes", title = "Status effect")
hist_status
top_interaction <- topTable(fit, coef = "infected.tb")
top_interaction
results_interaction <- topTable(fit, coef = "infected.tb",
                                number = nrow(counts), sort = "none")
ma_interaction <- ggplot(data.frame(Amean = fit$Amean, logFC = fit$coef[, "infected.tb"]),
                       aes(x = Amean, y = logFC)) +
  geom_point() +
  labs(x = "Average expression level", y = "Log fold change",
       title = "Interaction effect")
ma_interaction
hist_interaction <- ggplot(results_interaction, aes(x = P.Value)) +
  geom_histogram(binwidth = 0.01) +
  labs(x = "p-value", y = "Number of genes", title = "Interaction effect")
hist_interaction
ggsave("fig-limma-hist-interaction.eps", plot = hist_interaction,
       width = 7, height = 7, units = "in")
```

## Explore top hits

Boxplot function.

```{r boxplot-function}
plot_gene <- function(v, g) {
  # v - An EList object containing log2 counts per million
  # g - character vector of a single gene
  stopifnot(class(v) == "EList",
            is.character(g), length(g) == 1)
  library("tidyr")
  single_gene <- v$E[g, ]
  single_gene_long <- gather_(as.data.frame(single_gene),
                              gather_cols = colnames(single_gene))
  # For some reason, the argument value_col wouldn't work
  colnames(single_gene_long) <- "log2cpm"
  single_gene_long$sample <- rownames(single_gene_long)
  single_gene_long <- separate_(single_gene_long, col = "sample", sep = "-",
                                into = c("individual", "status", "treatment"))
  single_gene_long$status <- factor(single_gene_long$status, levels = c("contact", "tb"))
  single_gene_long$treatment <- factor(single_gene_long$treatment, levels = c("none", "infected"))
  ggplot(single_gene_long, aes(x = treatment, y = log2cpm, fill = status)) +
    geom_boxplot() +
    labs(title = g, x = "Treatment", y = expression("Expression level (" * log[2] * " cpm)"))
}
```

Infected.

```{r}
p_treatment_1 <- plot_gene(v, rownames(top_treatment)[1]) +
  ylim(3, 11) + theme(legend.position = "bottom")
p_treatment_2 <- plot_gene(v, rownames(top_treatment)[2]) +
  ylim(3, 11) + theme(legend.position = "bottom")
p_treatment_3 <- plot_gene(v, rownames(top_treatment)[3]) +
  ylim(3, 11) + theme(legend.position = "bottom")
p_treatment_4 <- plot_gene(v, rownames(top_treatment)[4]) +
  ylim(3, 11) + theme(legend.position = "bottom")
ggsave("fig-limma-treatment-hit-1.eps", plot = p_treatment_1,
       width = 7, height = 7, units = "in")
plot_grid(p_treatment_1, p_treatment_2, p_treatment_3, p_treatment_4)
```

Susceptibility status.

```{r}
plot_grid(plot_gene(v, rownames(top_status)[1]) +
            ylim(3, 11) + theme(legend.position = "bottom"),
          plot_gene(v, rownames(top_status)[2]) +
            ylim(3, 11) + theme(legend.position = "bottom"),
          plot_gene(v, rownames(top_status)[3]) +
            ylim(3, 11) + theme(legend.position = "bottom"),
          nrow = 1)
```

Interaction.

```{r}
plot_grid(plot_gene(v, rownames(top_interaction)[1]) +
            ylim(3, 11) + theme(legend.position = "bottom"),
          plot_gene(v, rownames(top_interaction)[2]) +
            ylim(3, 11) + theme(legend.position = "bottom"),
          plot_gene(v, rownames(top_interaction)[3]) +
            ylim(3, 11) + theme(legend.position = "bottom"),
          nrow = 1)
```

## Use ash for mutliple testing correction

Treatment effect.

```{r ash-treatment}
ash_treatment <- ash(betahat = fit$coefficients[, "infected"],
                     sebetahat = fit$stdev.unscaled[, "infected"] * sqrt(fit$s2.post),
                     df = fit$df.total)
class(ash_treatment)
names(ash_treatment)
sum(ash_treatment$svalue < .05)
hist(ash_treatment$svalue)
```

Susceptibility effect.

```{r ash-status}
ash_status <- ash(betahat = fit$coefficients[, "tb"],
                  sebetahat = fit$stdev.unscaled[, "tb"] * sqrt(fit$s2.post),
                  df = fit$df.total)
sum(ash_status$svalue < .05)
hist(ash_status$svalue)
```

Interaction effect.

```{r ash-interaction}
ash_interaction <- ash(betahat = fit$coefficients[, "infected.tb"],
                     sebetahat = fit$stdev.unscaled[, "infected.tb"] * sqrt(fit$s2.post),
                     df = fit$df.total)
sum(ash_interaction$svalue < .05)
hist(ash_interaction$svalue)
```

Venn diagram

```{r venn}
ash_treatment_de <- ash_treatment$svalue < .05
ash_status_de <- ash_status$svalue < .05
ash_interaction_de <- ash_interaction$svalue < .05
grid.newpage()
venn_rv <- draw.triple.venn(
  area1 = sum(ash_treatment_de),
  area2 = sum(ash_status_de),
  area3 = sum(ash_interaction_de),
  n12 = sum(ash_treatment_de & ash_status_de),
  n23 = sum(ash_status_de & ash_interaction_de),
  n13 = sum(ash_treatment_de & ash_interaction_de),
  n123 = sum(ash_treatment_de & ash_status_de & ash_interaction_de),
  category = c("Treatment", "Status", "Interaction"),
#   fill = c("darkolivegreen3", "cadetblue3", "darkorchid"),
  alpha = c(0.5, 0.5, 0.5), col = "black", cex = 2, cat.cex = 1.5,
  euler.d = FALSE, scaled = FALSE, ind = TRUE)
grid.draw(venn_rv)
```

## Save results of differential expression analysis

Save results from limma and ash.

```{r save-limma}
saveRDS(v, "../data/results-limma-voom.rds")
saveRDS(fit, "../data/results-limma-fit.rds")
saveRDS(ash_treatment, "../data/results-ash-treatment.rds")
saveRDS(ash_status, "../data/results-ash-status.rds")
saveRDS(ash_interaction, "../data/results-ash-interaction.rds")
```

## Create figure

```{r fig-limma}
theme_set(theme_bw(base_size = 8))
theme_update(panel.grid.minor.x = element_blank(),
             panel.grid.minor.y = element_blank(),
             panel.grid.major.x = element_blank(),
             panel.grid.major.y = element_blank())
fig_limma <- ggdraw() +
  draw_plot(ma_treatment,     0.00, 0.67, 0.33, 0.33) +
  draw_plot(ma_status,        0.00, 0.33, 0.33, 0.33) +
  draw_plot(ma_interaction,   0.00, 0.00, 0.33, 0.33) +
  draw_plot(hist_treatment,   0.33, 0.67, 0.33, 0.33) +
  draw_plot(hist_status,      0.33, 0.33, 0.33, 0.33) +
  draw_plot(hist_interaction, 0.33, 0.00, 0.33, 0.33) +
  draw_plot(p_treatment_1,    0.67, 0.82, 0.15, 0.15) +
  draw_plot(p_treatment_2,    0.67, 0.67, 0.15, 0.15) +
  draw_plot(p_treatment_3,    0.82, 0.67, 0.15, 0.15) +
  draw_plot(p_treatment_4,    0.82, 0.82, 0.15, 0.15) +
  draw_plot_label(c("A", "B", "C"), c(0.00, 0.00, 0.00),
                                    c(1.00, 0.67, 0.33), size = 15)
fig_limma
# Using cowplot ggsave, which by default does not use Dingbats font
ggsave("fig-limma.eps", plot = fig_limma, width = 8, height = 5, units = "in")
```
