---
title: "Normalization and filtering of bulkRNA Count data"
author: "Anthony Hung"
date: "2019-12-16"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Perform normalization of raw count data using different normalization methods and filter out lowly expressed genes

```{r load libraries and data}
library("gplots")
library("ggplot2")
library("reshape")
library("edgeR")
library("RColorBrewer")
library("scales")
library("cowplot")
library("dplyr")

# Load colors 
colors <- colorRampPalette(c(brewer.pal(9, "Blues")[1],brewer.pal(9, "Blues")[9]))(100)
pal <- c(brewer.pal(9, "Set1"), brewer.pal(8, "Set2"), brewer.pal(12, "Set3"))


# load in relabeled counts
raw_counts <- readRDS("data/raw_counts_relabeled.rds")
# Create DGEList object to allow for easier application of different normalization methods
raw_counts <- DGEList(raw_counts, group = colnames(raw_counts))

# load in reordered sample information
sampleinfo <- readRDS("data/Sample.info.RNAseq.reordered.csv")
```

# Correlation heatmap of Raw Data

```{r corr heatmap raw}
#Spearman
cors <- cor(raw_counts$counts, method="spearman", use="pairwise.complete.obs")

labels <- paste(sampleinfo$Individual, sampleinfo$treatment, sep=" ")
heatmap.2( cors, scale="none", col = colors, margins = c(12, 12), trace='none', denscol="white", labCol=labels, ColSideColors=pal[as.integer(as.factor(sampleinfo$Individual))], RowSideColors=pal[as.integer(as.factor(sampleinfo$treatment))+3], cexCol = 0.2 + 1/log10(15), cexRow = 0.2 + 1/log10(15))

#Pearson
cors <- cor(raw_counts$counts, method="pearson", use="pairwise.complete.obs")

labels <- paste(sampleinfo$Individual, sampleinfo$treatment, sep=" ")
heatmap.2( cors, scale="none", col = colors, margins = c(12, 12), trace='none', denscol="white", labCol=labels, ColSideColors=pal[as.integer(as.factor(sampleinfo$Individual))], RowSideColors=pal[as.integer(as.factor(sampleinfo$treatment))+3], cexCol = 0.2 + 1/log10(15), cexRow = 0.2 + 1/log10(15))
```

# Perform cpm normalization on all counts and plot density plots

```{r cpm normalization}
cpm <- cpm(raw_counts, log=TRUE)
head(cpm)

strained <- sampleinfo$treatment == "Strain"
unstrained <- sampleinfo$treatment == "Unstrain"
ind_1 <- sampleinfo$Individual == "NA18855 "
ind_2 <- sampleinfo$Individual == "NA18856 "
ind_3 <- sampleinfo$Individual == "NA19160 "


# Look at density plots for all individuals broken down by each treatment type
col = as.data.frame(pal[as.numeric(sampleinfo$Individual)])

plotDensities(cpm[,strained], col=col[strained, ], legend="topright")
plotDensities(cpm[,unstrained], col=col[unstrained, ], legend="topright")

# Look at density plots broken down by individual
col = as.data.frame(pal[as.numeric(sampleinfo$treatment)])

plotDensities(cpm[,ind_1], col=col[ind_1, ], legend="topright")
plotDensities(cpm[,ind_2], col=col[ind_2, ], legend="topright")
plotDensities(cpm[,ind_3], col=col[ind_3, ], legend="topright")
```

## Boxplots of cpm across samples

```{r boxplot cpm}
meltcpm <- melt(cpm)
names(meltcpm) <- c("gene", "sampleID", "cpm")
p <- ggplot(meltcpm, aes(factor(sampleID), cpm)) 
p + geom_boxplot() + theme(axis.text.x = element_text(angle = 90))
```

## Filtering for lowly expressed genes (avg log2CPM > 1.5 in at least 4 samples)

```{r filter CPM}
cutoff <- 1.5

keep <- rowSums(cpm > cutoff ) >=4

counts_cpm <- raw_counts[keep,]
filtered_cpm <- cpm(counts_cpm, log = T, normalized.lib.sizes = T)
dim(filtered_cpm)
```

## Boxplots of normalized+filtered counts across samples

```{r boxplot filt cpm}
melt_filt_cpm <- melt(filtered_cpm)
names(melt_filt_cpm) <- c("gene", "sampleID", "log2cpm")
p1 <- ggplot(melt_filt_cpm, aes(factor(sampleID), log2cpm)) 
p1 + geom_boxplot() + theme(axis.text.x = element_text(angle = 90))
```

## Correlation heatmap of log2CPM normalized samples

```{r corr heatmap CPM}
#Spearman
cors <- cor(filtered_cpm, method="spearman", use="pairwise.complete.obs")

labels <- paste(sampleinfo$Individual, sampleinfo$treatment, sep=" ")
heatmap.2( cors, scale="none", col = colors, margins = c(12, 12), trace='none', denscol="white", labCol=labels, ColSideColors=pal[as.integer(as.factor(sampleinfo$Individual))], RowSideColors=pal[as.integer(as.factor(sampleinfo$treatment))+3], cexCol = 0.2 + 1/log10(15), cexRow = 0.2 + 1/log10(15))

#Pearson
cors <- cor(filtered_cpm, method="pearson", use="pairwise.complete.obs")

labels <- paste(sampleinfo$Individual, sampleinfo$treatment, sep=" ")
heatmap.2( cors, scale="none", col = colors, margins = c(12, 12), trace='none', denscol="white", labCol=labels, ColSideColors=pal[as.integer(as.factor(sampleinfo$Individual))], RowSideColors=pal[as.integer(as.factor(sampleinfo$treatment))+3], cexCol = 0.2 + 1/log10(15), cexRow = 0.2 + 1/log10(15))
```


# TMM Normalization

```{r TMM normalization}
TMM <- calcNormFactors(raw_counts, method = "TMM")
TMM <- cpm(TMM, log=TRUE)
head(TMM)

strained <- sampleinfo$treatment == "Strain"
unstrained <- sampleinfo$treatment == "Unstrain"
ind_1 <- sampleinfo$Individual == "NA18855 "
ind_2 <- sampleinfo$Individual == "NA18856 "
ind_3 <- sampleinfo$Individual == "NA19160 "


# Look at density plots for all individuals broken down by each treatment type
col = as.data.frame(pal[as.numeric(sampleinfo$Individual)])

plotDensities(TMM[,strained], col=col[strained, ], legend="topright")
plotDensities(TMM[,unstrained], col=col[unstrained, ], legend="topright")

# Look at density plots broken down by individual
col = as.data.frame(pal[as.numeric(sampleinfo$treatment)])

plotDensities(TMM[,ind_1], col=col[ind_1, ], legend="topright")
plotDensities(TMM[,ind_2], col=col[ind_2, ], legend="topright")
plotDensities(TMM[,ind_3], col=col[ind_3, ], legend="topright")
```


## Boxplots of TMM across samples

```{r boxplot TMM}
meltTMM <- melt(TMM)
names(meltTMM) <- c("gene", "sampleID", "TMM")
p <- ggplot(meltTMM, aes(factor(sampleID), TMM)) 
p + geom_boxplot() + theme(axis.text.x = element_text(angle = 90))
```

## Filtering for lowly expressed genes (avg log2TMM > 1.5 in at least 4 samples)

```{r filter TMM}
cutoff <- 1.5

keep <- rowSums( TMM > cutoff ) >=4

counts_TMM <- raw_counts[keep,]
filtered_TMM <- cpm(counts_TMM, log = T, normalized.lib.sizes = T)
dim(filtered_TMM)
```

## Boxplots of normalized+filtered counts across samples

```{r boxplot filt TMM}
melt_filt_TMM <- melt(filtered_TMM)
names(melt_filt_TMM) <- c("gene", "sampleID", "log2TMM")
p1 <- ggplot(melt_filt_TMM, aes(factor(sampleID), log2TMM)) 
p1 + geom_boxplot() + theme(axis.text.x = element_text(angle = 90))
```

## Correlation heatmap of log2TMM normalized samples

```{r corr heatmap TMM}
#Spearman
cors <- cor(filtered_TMM, method="spearman", use="pairwise.complete.obs")

labels <- paste(sampleinfo$Individual, sampleinfo$treatment, sep=" ")
heatmap.2( cors, scale="none", col = colors, margins = c(12, 12), trace='none', denscol="white", labCol=labels, ColSideColors=pal[as.integer(as.factor(sampleinfo$Individual))], RowSideColors=pal[as.integer(as.factor(sampleinfo$treatment))+3], cexCol = 0.2 + 1/log10(15), cexRow = 0.2 + 1/log10(15))


#pearson
cors <- cor(filtered_TMM, method="pearson", use="pairwise.complete.obs")

labels <- paste(sampleinfo$Individual, sampleinfo$treatment, sep=" ")
heatmap.2( cors, scale="none", col = colors, margins = c(12, 12), trace='none', denscol="white", labCol=labels, ColSideColors=pal[as.integer(as.factor(sampleinfo$Individual))], RowSideColors=pal[as.integer(as.factor(sampleinfo$treatment))+3], cexCol = 0.2 + 1/log10(15), cexRow = 0.2 + 1/log10(15))
```




# RLE Normalization

```{r RLE normalization}
RLE <- calcNormFactors(raw_counts, method = "RLE")
RLE <- cpm(RLE, log=TRUE)
head(RLE)

strained <- sampleinfo$treatment == "Strain"
unstrained <- sampleinfo$treatment == "Unstrain"
ind_1 <- sampleinfo$Individual == "NA18855 "
ind_2 <- sampleinfo$Individual == "NA18856 "
ind_3 <- sampleinfo$Individual == "NA19160 "


# Look at density plots for all individuals broken down by each treatment type
col = as.data.frame(pal[as.numeric(sampleinfo$Individual)])

plotDensities(RLE[,strained], col=col[strained, ], legend="topright")
plotDensities(RLE[,unstrained], col=col[unstrained, ], legend="topright")

# Look at density plots broken down by individual
col = as.data.frame(pal[as.numeric(sampleinfo$treatment)])

plotDensities(RLE[,ind_1], col=col[ind_1, ], legend="topright")
plotDensities(RLE[,ind_2], col=col[ind_2, ], legend="topright")
plotDensities(RLE[,ind_3], col=col[ind_3, ], legend="topright")
```


## Boxplots of RLE across samples

```{r boxplot RLE}
meltRLE <- melt(RLE)
names(meltRLE) <- c("gene", "sampleID", "RLE")
p <- ggplot(meltRLE, aes(factor(sampleID), RLE)) 
p + geom_boxplot() + theme(axis.text.x = element_text(angle = 90))
```

## Filtering for lowly expressed genes (avg log2RLE > 1.5 in at least 4 samples)

```{r filter RLE}
cutoff <- 1.5

keep <- rowSums( RLE > cutoff ) >=4

counts_RLE <- raw_counts[keep,]
filtered_RLE <- cpm(counts_RLE, log = T, normalized.lib.sizes = T)
dim(filtered_RLE)
```

## Boxplots of normalized+filtered counts across samples

```{r boxplot filt RLE}
melt_filt_RLE <- melt(filtered_RLE)
names(melt_filt_RLE) <- c("gene", "sampleID", "log2RLE")
p1 <- ggplot(melt_filt_RLE, aes(factor(sampleID), log2RLE)) 
p1 + geom_boxplot() + theme(axis.text.x = element_text(angle = 90))
```

## Correlation heatmap of log2RLE normalized samples

```{r corr heatmap RLE}
#Spearman
cors <- cor(filtered_RLE, method="spearman", use="pairwise.complete.obs")

labels <- paste(sampleinfo$Individual, sampleinfo$treatment, sep=" ")
heatmap.2( cors, scale="none", col = colors, margins = c(12, 12), trace='none', denscol="white", labCol=labels, ColSideColors=pal[as.integer(as.factor(sampleinfo$Individual))], RowSideColors=pal[as.integer(as.factor(sampleinfo$treatment))+3], cexCol = 0.2 + 1/log10(15), cexRow = 0.2 + 1/log10(15))


#pearson
cors <- cor(filtered_RLE, method="pearson", use="pairwise.complete.obs")

labels <- paste(sampleinfo$Individual, sampleinfo$treatment, sep=" ")
heatmap.2( cors, scale="none", col = colors, margins = c(12, 12), trace='none', denscol="white", labCol=labels, ColSideColors=pal[as.integer(as.factor(sampleinfo$Individual))], RowSideColors=pal[as.integer(as.factor(sampleinfo$treatment))+3], cexCol = 0.2 + 1/log10(15), cexRow = 0.2 + 1/log10(15))
```


# upperquartile Normalization

```{r upperquartile normalization}
upperquartile <- calcNormFactors(raw_counts, method = "upperquartile")
upperquartile <- cpm(upperquartile, log=TRUE)
head(upperquartile)

strained <- sampleinfo$treatment == "Strain"
unstrained <- sampleinfo$treatment == "Unstrain"
ind_1 <- sampleinfo$Individual == "NA18855 "
ind_2 <- sampleinfo$Individual == "NA18856 "
ind_3 <- sampleinfo$Individual == "NA19160 "


# Look at density plots for all individuals broken down by each treatment type
col = as.data.frame(pal[as.numeric(sampleinfo$Individual)])

plotDensities(upperquartile[,strained], col=col[strained, ], legend="topright")
plotDensities(upperquartile[,unstrained], col=col[unstrained, ], legend="topright")

# Look at density plots broken down by individual
col = as.data.frame(pal[as.numeric(sampleinfo$treatment)])

plotDensities(upperquartile[,ind_1], col=col[ind_1, ], legend="topright")
plotDensities(upperquartile[,ind_2], col=col[ind_2, ], legend="topright")
plotDensities(upperquartile[,ind_3], col=col[ind_3, ], legend="topright")
```


## Boxplots of upperquartile across samples

```{r boxplot upperquartile}
meltupperquartile <- melt(upperquartile)
names(meltupperquartile) <- c("gene", "sampleID", "upperquartile")
p <- ggplot(meltupperquartile, aes(factor(sampleID), upperquartile)) 
p + geom_boxplot() + theme(axis.text.x = element_text(angle = 90))
```

## Filtering for lowly expressed genes (avg log2upperquartile > 1.5 in at least 4 samples)

```{r filter upperquartile}
cutoff <- 1.5

keep <- rowSums( upperquartile > cutoff ) >=4

counts_upperquartile <- raw_counts[keep,]
filtered_upperquartile <- cpm(counts_upperquartile, log = T, normalized.lib.sizes = T)
dim(filtered_upperquartile)
```

## Boxplots of normalized+filtered counts across samples

```{r boxplot filt upperquartile}
melt_filt_upperquartile <- melt(filtered_upperquartile)
names(melt_filt_upperquartile) <- c("gene", "sampleID", "log2upperquartile")
p1 <- ggplot(melt_filt_upperquartile, aes(factor(sampleID), log2upperquartile)) 
p1 + geom_boxplot() + theme(axis.text.x = element_text(angle = 90))
```

## Correlation heatmap of log2upperquartile normalized samples

```{r corr heatmap upperquartile}
#Spearman
cors <- cor(filtered_upperquartile, method="spearman", use="pairwise.complete.obs")

labels <- paste(sampleinfo$Individual, sampleinfo$treatment, sep=" ")
heatmap.2( cors, scale="none", col = colors, margins = c(12, 12), trace='none', denscol="white", labCol=labels, ColSideColors=pal[as.integer(as.factor(sampleinfo$Individual))], RowSideColors=pal[as.integer(as.factor(sampleinfo$treatment))+3], cexCol = 0.2 + 1/log10(15), cexRow = 0.2 + 1/log10(15))


#pearson
cors <- cor(filtered_upperquartile, method="pearson", use="pairwise.complete.obs")

labels <- paste(sampleinfo$Individual, sampleinfo$treatment, sep=" ")
heatmap.2( cors, scale="none", col = colors, margins = c(12, 12), trace='none', denscol="white", labCol=labels, ColSideColors=pal[as.integer(as.factor(sampleinfo$Individual))], RowSideColors=pal[as.integer(as.factor(sampleinfo$treatment))+3], cexCol = 0.2 + 1/log10(15), cexRow = 0.2 + 1/log10(15))
```




## Save normalized/filtered count matrix and filtered count matrix

```{r save}
saveRDS(counts_cpm, "data/filtered_counts.rds")
saveRDS(filtered_cpm, "data/norm_filtered_counts.rds")
```
